---
title: "breastCancerExploration"
author: "Tony Song"
format: html
editor: visual
---

## Breast Cancer Dataset

Load necessary libraries

```{r}
library(utils)
library(ggcorrplot)
library(ggplot2)
library(dplyr)
library(tidyr)
library(reshape2)
library(gridExtra)
library(paletteer)
```

Download data then upload to R.

```{r}
setwd("C:/Users/Tony/OneDrive/Documents/tonyR/DSRP-2024-Khevna/tonyWork")
data <- read.csv("../data/breast_cancer_classification_data.csv")
```

Let's look at some of the elements in this dataset

```{r}
head(data)
```

Now, in order to manipulate the dataset well, we need to get the dimensions

```{r}
dim(data)
```

This shows that there are 569 rows and 33 columns. Now, we need to know what these column names are and what data type they are.

```{r}
str(data)
```

We can see that other than the diagnosis, all of the variables are numerical.

First, let's explore a bit of our categorical variable: diagnosis. A diagnosis of M means that the cell is malignant while a diagnosis of B means that it is benign. We can plot this with a multitude of graphs

```{r}
ggplot(data,aes(x=diagnosis,fill=diagnosis)) + geom_bar() + scale_fill_paletteer_d("ggthemes::Tableau_10")

ggplot(data,aes(x="",fill=diagnosis)) + geom_bar() + coord_polar(start=0,"y") + theme_void() + scale_fill_paletteer_d("ggthemes::Tableau_10")
```
